Existence of Forward Simulations for Particular Data Structures

### 2017-01-13 draftCAV/commit.tex
started the stack implementation

\section{Relaxation for the Data Structures Without Fixed Remove Linearization Points}
We have observed that some implementations (like time-stamped stack) do not have fixed remove (pop) linearization points that will correspond to $lin(pop,e,k)$ where $e$ could be a data value or $\texttt{EMPTY}$. However, we observe that, these implementations contain some points in their pop methods that logically removes the element from the pool. We call them commit points. If a method of a library has a fixed linearization  point, it is also a commit point. In this sense, commit points are weaker versions of fixed linearization points. Fixed linearization point of a pop preserves the following properties:
\begin{itemize}
\item If a $ret$ comes before a linearization point in the concrete execution, this order is preserved in the linearization of this execution.
\item If a $lin$ pop comes before another $lin$ pop in the concrete execution, this order is preserved in the linearization of this execution.
\item If a $lin$ pop comes before an $inv$ in the concrete execution, this order is preserved in the linearization of this execution. 
\end{itemize} 
A commit point is weaker than a fixed linearization point in the sense that it does not need to satisfy the first and the second conditions.

>>>>

%!TEX root = draft.tex
\section{Stacks With Fixed Pop Commit Points}

Differently from queues, there are implementations of concurrent stacks where the linearization points of the remove/pop operations are not fixed, e.g., Time-stamped Stack~\cite{}. We define an abstract implementation which simulates such implementations provided that the point in time where the return value of a pop operation is determined corresponds to a fixed action. To demonstrate the use of this abstract implementation we provide a correctness proof for a simplified version of Time-Stamped Stack that preserves its most complex behaviors.

\subsection{Abstract stack implementation}

Let $Com(pop)$ be the set of actions $com(pop,d,k)$ with $d\in\<Vals>$ and $k\in\<Ops>$, called \emph{commit points}, representing actions of a pop operation where its return value is determined. We define an abstract implementation $AbsS$ over alphabet $C\cup R\cup Com(pop)$ that essentially, similarly to $AbsQ$, maintains the happens-before order of the push operations whose value has not been yet removed. Since the commit points are not necessarily linearization points, the pop operations are treated differently. Intuitively, each pop operation starts by taking a snapshot of the completed push operations which are maximal in the happens-before order and continuously tracks the push operations which are overlapping with it. The commit point is enabled only if the return value corresponds to one of the push operations in the initial snapshot, or to a push happening earlier when values from the initial snapshot are removed by other pop operations, or to one of the push operations that overlaps with the pop.

The states of $AbsS$ are tuples $\tup{O,<,\ell,rval,cp,comp,pend}$ where $O\subseteq \<Ops>$ represents a set of operation identifiers (of previously invoked push operations), $<\subseteq O\times O$ is a strict partial order over $O$, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier in $O$ with a value and a flag that records whether the operation is pending or completed, $coPush:\<Ops> ~> 2^O$ records push operations that were maximal when a pop started or happening earlier provided that the values of all the push happening later have been removed, and $pePush: \<Ops> ~> 2^O$ records push operations overlapping with a pop.
The initial state has all these components set to $\emptyset$ and the transition relation $->$ is defined in Figure~\ref{fig:transitions:AbsS}.

\begin{figure} [t]
{\footnotesize
  \centering
  \begin{mathpar}
    \inferrule[call-push]{
      k\not\in dom(cp) \\ 
    }{
      O,<,\ell,rval,cp,comp,pend 
      \xrightarrow{inv(push,d,k)} \\
      O\cup\{k\},<\cup \{(k',k): \ell_2(k')={\tt COMP}\},\ell[k\mapsto (d,{\tt PEND},\emptyset)],rval,cp[k\mapsto 1],comp,pend[\forall k'.\ k'\mapsto pend(k')\cup \{k\}]
    }\hspace{5mm}

    \inferrule[call-pop]{
      k\not\in dom(cp) \\
    }{
      O,<,\ell,rval,cp,comp,pend
      \xrightarrow{inv(pop,k)} \\
      O,<,\ell,rval,cp[k\mapsto 1],comp[k\mapsto maxCompl(<)],pend[k\mapsto \{k':\ell_2(k')={\tt PEND}\}]
    }\hspace{5mm}

    \inferrule[ret-pop]{
       cp(k) = 2 \\
       rval(k)=d  
    }{
      O,<,\ell,rval,cp,comp,pend
      \xrightarrow{ret(pop,d,k)}
      O,<,\ell,rval,cp[k\mapsto 3],comp,pend
    }\hspace{5mm}

    \inferrule[ret-push1]{
      cp(k) = 1 \\
      k \in O \\
      \ell(k) = (d,{\tt PEND},X) 
    }{
      O,<,\ell,rval,cp,comp,pend
      \xrightarrow{ret(push,k)}
      O,<,\ell[k\mapsto (d,{\tt COMP},X)],rval,cp[k\mapsto 2],comp,pend
    }\hspace{5mm}

    \inferrule[ret-push2]{
      cp(k) = 1 \\
      k \not\in O 
    }{
      O,<,\ell,rval,cp,comp,pend
      \xrightarrow{ret(push,k)}
      O,<,\ell,rval,cp[k\mapsto 2],comp,pend
    }\hspace{5mm}

    \inferrule[lin-pop1]{
       cp(k) = 1 \\
       d\neq{\tt EMPTY} \\
       \exists k'\in O.\ k'\in comp(k)\cup pend(k)\land \ell_1(k')=d
    }{
      O,<,\ell,rval,cp,comp,pend
      \xrightarrow{lin(pop,d,k)} \\
      O\setminus \{k\},<\uparrow k,\ell,rval[k\mapsto d],cp[k\mapsto 2],comp[\forall k''.\ k'\in comp(k'') \Rightarrow k'' \mapsto (comp(k'')\cup pred_{<}(k'))\setminus\{k'\}],pend
    }\hspace{5mm}

    \inferrule[lin-deq2]{
       cp(k) = 1 \\
       comp(k)=\emptyset
    }{
      O,<,\ell,rval,cp,comp,pend
      \xrightarrow{lin(deq,{\tt EMPTY},k)}
      O,<,\ell,rval[k\mapsto {\tt EMPTY}],cp[k\mapsto 2],comp,pend
    }\hspace{5mm}

      \end{mathpar}
  }
 \vspace{-5mm}
  \caption{The transition relation of $AbsQ$.}
  \label{fig:transitions:AbsS}
\vspace{-6mm}
\end{figure}

\subsection{Other}

We have observed that some implementations (like time-stamped stack) do not have fixed remove (pop) linearization points that will correspond to $lin(pop,e,k)$ where $e$ could be a data value or $\texttt{EMPTY}$. However, we observe that, these implementations contain some points in their pop methods that logically removes the element from the pool. We call them commit points. If a method of a library has a fixed linearization  point, it is also a commit point. In this sense, commit points are weaker versions of fixed linearization points. Fixed linearization point of a pop preserves the following properties:
\begin{itemize}
\item If a $ret$ comes before a linearization point in the concrete execution, this order is preserved in the linearization of this execution.
\item If a $lin$ pop comes before another $lin$ pop in the concrete execution, this order is preserved in the linearization of this execution.
\item If a $lin$ pop comes before an $inv$ in the concrete execution, this order is preserved in the linearization of this execution. 
\end{itemize} 
A commit point is weaker than a fixed linearization point in the sense that it does not need to satisfy the first and the second conditions.

### 2017-01-13 draftCAV/queue.tex
started the stack implementation

The states of $AbsQ$ are tuples $\tup{O,<,\ell,rval,cp}$ where $O\subseteq \<Ops>$ represents a set of operation identifiers (of previously invoked enqueue operations), $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a flag that records whether the operation is pending or completed, $rval:\<Ops> ~> \<Vals>$ records the return value of a pending dequeue fixed at its linearization point, and $cp:\<Ops> ~> \<Nats>$ records the control point of every pending (enqueue or dequeue) operation.
The initial state is $\tup{\emptyset,\emptyset,\emptyset,\emptyset,\emptyset}$ and the transition relation $->$ is defined in Figure~\ref{fig:transitions:AbsQ}.

@@

>>>>

The states of $AbsQ$ are tuples $\tup{O,<,\ell,rval,cp}$ where $O\subseteq \<Ops>$ represents a set of operation identifiers (of previously invoked enqueue operations), $<\subseteq O\times O$ is a strict partial order, $\ell: O -> \<Vals>\times\{\tt{PEND,\tt{COMP}}\}$ labels every identifier with a value and a flag that records whether the operation is pending or completed, $rval:\<Ops> ~> \<Vals>$ records the return value of a pending dequeue fixed at its linearization point, and $cp:\<Ops> ~> \<Nats>$ records the control point of every pending (enqueue or dequeue) operation.
The initial state has all these components set to $\emptyset$ and the transition relation $->$ is defined in Figure~\ref{fig:transitions:AbsQ}.

@@

TODO NEEDS DATA INDEPENDENCE FOR THE LINEARIZATION POINT TRANSITIONS TO BE DETERMINISTIC
